tl20-05_berghei_pir_GeneSetEnrichmentQusage

### 2021-10-04 chapter3_results_bergheiPir/Manuscript_berghei_pir_DESEQ.Rmd
Add colored labels to Qusage heatmap

```{r load_packages}
library('dplyr')
library('readr')
library('readxl')
library('writexl')
library('data.table')
library('tibble')
library('ggplot2')
library('stringr')
library('RColorBrewer')
library('DESeq2')
library('apeglm')
library('ashr')
```

@@

```{r lfcShrink}
shrunk.list <- lapply(comparisons, 
                      function(comparison){
                        res <- results(
                          dds.deseq,
                          contrast = c("stage",comparison[1],comparison[2])
                        )
                        shrunk <- lfcShrink(dds.deseq, 
                                            res = res,
                                            type = "ashr") %>% 
                          as.data.frame %>% 
                          rownames_to_column(var = 'Geneid') %>% 
                          mutate(comp = paste0(comparison[1],'_',comparison[2])) %>% 
                          filter(Geneid %in% pir_id,
                                 !is.na(padj))
                        return(shrunk)
                      })
names(shrunk.list) <- unlist(lapply(comparisons, 
                                    function(comparison) paste0(comparison[1],'_',comparison[2])))
sig_pirs <- bind_rows(shrunk.list) %>% 
  mutate(subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
  arrange(subfam, Geneid) %>% 
  filter(padj < 0.05)
sig_pirs_order <- unique(sig_pirs$Geneid)
sig_pirs_cast <- dcast(sig_pirs,
                       comp ~ Geneid,
                       value.var = 'log2FoldChange') %>% 
  column_to_rownames(var = 'comp')
sig_pirs_cast <- sig_pirs_cast[ordering(rownames(sig_pirs_cast),
                                        c(r_order_vector,
                                          'Gam.Male_Asex.Mixed',
                                          'Gam.Male_Liv.Meroz')),
                               sig_pirs_order]

@@

```{r heatmap}
#Add a new-line to the S6 label so that it doesn't overlap with others
pir_column_split <- gsub(
  gsub(pir_clade$clade[match(colnames(sig_pirs_cast), pir_clade$Geneid)],
       pattern = 'ancestral',
       replacement = 'a'),
  pattern = '(S6|L2)',
  replacement = '\\1\n'
)

png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_inidvidualGenes.png",
    width = 7,
    height = 5,
    units = "in",
    res = 330)
Heatmap(sig_pirs_cast,
        name = 'logFoldChange',
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        # rect_gp = gpar(col = "white", lwd = 1),
        border_gp = gpar(col = "black", lty = 1),
        column_title_gp = gpar(fontsize = 12),
        na_col = 'white',
        show_column_names = FALSE,
        column_split = pir_column_split,
        heatmap_height = unit(10, "cm"),
        #Shrink heatmap a little so that the S6 label is visible
        column_gap = unit(2, 'mm')
)
dev.off()
```

```{r save_stats_df}
deseq_stats <- bind_rows(shrunk.list) %>% 
  mutate(subfam = pir_clade$clade[match(Geneid, pir_clade$Geneid)]) %>% 
  arrange(subfam, Geneid)

>>>>

```{r load_packages}
library('dplyr')
library('readr')
library('readxl')
library('writexl')
library('data.table')
library('tibble')
library('ggplot2')
library('stringr')
library('RColorBrewer')
library('DESeq2')
library('apeglm')
library('ashr')
library('ComplexHeatmap')
library('gridtext') #For heatmap labels
```

@@

```{r lfcShrink}
shrunk.list <- lapply(comparisons, 
                      function(comparison){
                        res <- results(
                          dds.deseq,
                          contrast = c("stage",comparison[1],comparison[2])
                        )
                        shrunk <- lfcShrink(dds.deseq, 
                                            res = res,
                                            type = "ashr") %>% 
                          as.data.frame %>% 
                          rownames_to_column(var = 'Geneid') %>% 
                          mutate(comp = paste0(comparison[1],'_',comparison[2])) %>% 
                          filter(Geneid %in% pir_geneset$pirs,
                                 !is.na(padj))
                        return(shrunk)
                      })
names(shrunk.list) <- unlist(lapply(comparisons, 
                                    function(comparison) paste0(comparison[1],'_',comparison[2])))
sig_pirs <- bind_rows(shrunk.list) %>% 
  mutate(subfam = pir_info$clade[match(Geneid, pir_info$Geneid)]) %>% 
  arrange(subfam, Geneid) %>% 
  filter(padj < 0.05)
sig_pirs_order <- unique(sig_pirs$Geneid)
sig_pirs_cast <- dcast(sig_pirs,
                       comp ~ Geneid,
                       value.var = 'log2FoldChange') %>% 
  column_to_rownames(var = 'comp')
sig_pirs_cast <- sig_pirs_cast[ordering(rownames(sig_pirs_cast),
                                        c(r_order_vector,
                                          'Gam.Male_Asex.Mixed',
                                          'Gam.Male_Liv.Meroz')),
                               sig_pirs_order]

@@

```{r heatmap}
#Add a new-line to the S6 label so that it doesn't overlap with others
pir_column_split <- gsub(
  gsub(pir_info$clade[match(colnames(sig_pirs_cast), pir_info$Geneid)],
       pattern = 'ancestral',
       replacement = 'a'),
  pattern = '(S6|L2)',
  replacement = '\\1\n'
)

#Make nice colour labels
row_labels <- gt_render(
  unlist(
    lapply(rownames(sig_pirs_cast), 
           function(comparison) {
             comp_sep <- unlist(
               strsplit(comparison, 
                        split = '_')
             )
             txt.colour <- sapply(comp_sep,
                                  function(stage_name){
                                    case_when(
                                      grepl(stage_name, pattern = 'Asex') ~ 'coral1' ,
                                      grepl(stage_name, pattern = 'Liv') ~ '#BD65B7',
                                      grepl(stage_name, pattern = 'Gam') ~ '#6998F4',
                                      grepl(stage_name, pattern  = 'Bld|Sporo|Ook') ~ '#629C57')
                                  })
             paste0("<span style='color:",
                    txt.colour[1],
                    "'>**",
                    comp_sep[1],
                    "** </span><span style='color:",
                    txt.colour[2],
                    "'>**",
                    comp_sep[2],
                    "**</span>")
           })
  )
)

png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_inidvidualGenes.png",
    width = 7,
    height = 5,
    units = "in",
    res = 330)
Heatmap(sig_pirs_cast,
        name = 'logFoldChange',
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        row_labels = row_labels,
        # rect_gp = gpar(col = "white", lwd = 1),
        border_gp = gpar(col = "black", lty = 1),
        column_title_gp = gpar(fontsize = 12),
        na_col = 'white',
        show_column_names = FALSE,
        column_split = pir_column_split,
        heatmap_height = unit(10, "cm"),
        #Shrink heatmap a little so that the S6 label is visible
        column_gap = unit(2, 'mm')
)
dev.off()
```

```{r save_stats_df}
deseq_stats <- bind_rows(shrunk.list) %>% 
  mutate(subfam = pir_info$clade[match(Geneid, pir_info$Geneid)]) %>% 
  arrange(subfam, Geneid)

### 2021-10-04 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Remove Asex.Ring v Mix comparison Qusage
Add colored labels to Qusage heatmap

#Remove some comparisons
comparisons <- comparisons[!grepl(comparisons, pattern = 'Asex\\.Mixed-Asex\\.Ring')]

@@

png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_geneGroups.png",
    width = 7,
    height = 5,
    units = "in",
    res = 330)
Heatmap(all_stats_df_cast,
        name = 'logFoldChange',
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        rect_gp = gpar(col = "white", lwd = 2),
        cell_fun = function(j, i, x, y, width, height, fill){
          stat_df <- all_stats_df[all_stats_df$contrast == rownames(all_stats_df_cast)[i],]
          stat_df <- stat_df[stat_df$pathway.name == colnames(all_stats_df_cast)[j],]
          fdr_val <- stat_df$FDR
          logfold <- stat_df$log.fold.change
          if(fdr_val <= 0.05){
            grid.text(
              sprintf("%.3f", #One significant figure
                      fdr_val),
              x,
              y,
              gp = gpar(fontsize = 10,
                        fontface = 'bold',
                        col = ifelse(logfold > 2, 'white', 'black'))
            )
          }
        })
dev.off()
```

```{r save_xlsx}
write_xlsx(all_stats_df,
           path = 'SupplementaryInformation_Stats_Qusage_berghei.xlsx')
write.csv(all_stats_df,
          'SupplementaryInformation_Stats_Qusage_berghei.csv')
```

>>>>

#Remove some comparisons
comparisons <- comparisons[!grepl(comparisons, pattern = 'Asex\\.Ring-Asex\\.Mixed')]

@@

#Make nice colour labels
row_labels <- gt_render(
  unlist(
    lapply(rownames(all_stats_df_cast), 
           function(comparison) {
             comp_sep <- unlist(
               strsplit(comparison, 
                        split = '-')
             )
             txt.colour <- sapply(comp_sep,
                                  function(stage_name){
                                    case_when(
                                      grepl(stage_name, pattern = 'Asex') ~ 'coral1' ,
                                      grepl(stage_name, pattern = 'Liv') ~ '#BD65B7',
                                      grepl(stage_name, pattern = 'Gam') ~ '#6998F4',
                                      grepl(stage_name, pattern  = 'Bld|Sporo|Ook') ~ '#629C57')
                                  })
             paste0("<span style='color:",
                    txt.colour[1],
                    "'>**",
                    comp_sep[1],
                    "** </span><span style='color:",
                    txt.colour[2],
                    "'>**",
                    comp_sep[2],
                    "**</span>")
           })
  )
)

png(filename = "plots/thesis_chapter3_figure_statsCompHeatmap_geneGroups.png",
    width = 7,
    height = 5,
    units = "in",
    res = 330)
Heatmap(all_stats_df_cast,
        name = 'logFoldChange',
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        row_labels = row_labels,
        rect_gp = gpar(col = "white", lwd = 2),
        cell_fun = function(j, i, x, y, width, height, fill){
          stat_df <- all_stats_df[all_stats_df$contrast == rownames(all_stats_df_cast)[i],]
          stat_df <- stat_df[stat_df$pathway.name == colnames(all_stats_df_cast)[j],]
          fdr_val <- stat_df$FDR
          logfold <- stat_df$log.fold.change
          if(fdr_val <= 0.05){
            grid.text(
              sprintf("%.3f", #One significant figure
                      fdr_val),
              x,
              y,
              gp = gpar(fontsize = 10,
                        fontface = 'bold',
                        col = ifelse(logfold > 2, 'white', 'black'))
            )
          }
        })
dev.off()
```

```{r save_xlsx}
write_xlsx(all_stats_df,
           path = 'SupplementaryInformation_Stats_Qusage_berghei.xlsx')
write.csv(all_stats_df,
          'SupplementaryInformation_Stats_Qusage_berghei.csv')
# all_stats_df <- read.csv('SupplementaryInformation_Stats_Qusage_berghei.csv',
#                          row.names = 1)
```
